---
title: "coffee producer"
author: "Abosede"
date: "04/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("ggplot2")
install.packages("ggthemes")
install.packages("dplyr")
library(ggplot2)
library(ggthemes)
library(scales)
library(dplyr)
```
##Introduction
##Analyses of Coffee producer data in three region, Africa and the middle East, Americas, Asia and Pacific,
##The data contains information on conventional Production, Organic Production and Area of land covered by each
#The code below is to detect and list out outliers in the coffee producer organization data,
```{r}
rm(list = ls())
options(scipen = 999) 
```

```{r}
# get the working directory or load R-project.
getwd()
```
```{r}
# Set the working directory or load R-project.
setwd("C:/users/admin/oneDrive/Desktop/Fairtrade")
```
```{r}
mydata<- read.csv("Producerdata.csv")
head(mydata)
summary(mydata)
```
##The result above are the initial description ,head function shows  8 variable with 526 observations. and shows that the variables are in character format. For quantitative analysis to be performed the variables have to be converted to numeric

## Converting variables from character to numeric
```{r}
mydata$Area.total<- as.numeric(as.character(mydata$Area.total))
mydata$Area.conventional<- as.numeric(as.character(mydata$Area.conventional))
mydata$Area.organic<- as.numeric(as.character(mydata$Area.organic))
mydata$Production.conventional<- as.numeric(as.character(mydata$Production.conventional))
mydata$Production.Organic<- as.numeric(as.character(mydata$Production.Organic))
mydata$Production.total<- as.numeric(as.character(mydata$Production.total))
```
```{r}
summary(mydata)
```

#Determining Outliers
##Outliers are values that are significantly different or distant  from other values
##It is determined using boxplot and stats function
#The box plot show five number summary of a dataset(minimum score, first quartile,median, third quatile and maximum score)
## the box plot also shows the distribution of the data 
##Any value outside the whiskers of the box plot is referred to as outliers
```{r}
boxplot(mydata$Production.conventional,
        ylab="Production conventional",col = rainbow(2),notch = TRUE,
        main="Boxplot of conventional production")

```

```{r}
boxplot(mydata$Production.Organic,
        ylab="Production.Orgnic",col = rainbow(2),notch = TRUE,
        main="Boxplot of organic production")
```

```{r}
boxplot(mydata$Production.total,
        ylab="Production.total",col = rainbow(2),notch = TRUE,
        main="Boxplot of total production")
```



```{r}
boxplot.stats(mydata$Production.conventional)$out

```
```{r}
boxplot.stats(mydata$Production.Organic)$out
```
```{r}
boxplot.stats(mydata$Production.total)$out
```



## The boxplot and the stats result shows that their is no outliers in Production organic and total production but their are two outliers in production conventional
##extracting the row number corresponding to the outliers
```{r}
out<-boxplot.stats(mydata$Production.conventional)$out
out_ind<-which(mydata$Production.conventional%in%c(out))
out_ind
```
```{r}
mydata[out_ind,]
```
```{r}
boxplot(mydata$Production.conventional,
        ylab="Production conventional",col = rainbow(2),notch = TRUE)

mtext(paste("Outliers:",paste(out,collapse = ",")))
```
##Boxplot was use to detect outliers using information such as  first quartile, third quartile and interquartile range
##stats function was used to list out the outliers
#formula
#IQR=Q3-Q1
#I=[Q1-1.5*IQR;Q3+1.5*IQR]
#numbers below the high and low level are referred to as outliers


##Question 2
##The first aggegrate is done without considering outliers
## Aggregated statistics on production and area of coffee per region
## The analysis was done using stacked bar chart

##Aggregation of variables by Region
```{r}
Aggregate<-mydata%>%
  group_by(Region)%>%
  summarise(across(c(Area.total,Area.conventional,Area.organic,
                     Production.conventional,Production.Organic,Production.total),sum,na.rm=TRUE))
Aggregate
```
```{r}
Areatotal$AreaOrganic<-mydata%>%
  group_by(Region)%>%
  summarize(sum_yield=sum(Area.organic,na.rm = TRUE))
```
##Showing aggregate using stacked barchart
```{r}
ggplot(Aggregate, aes(x=Area.total,fill=Region))+geom_bar()
```
```{r}
```
```{r}
```


##Eliminating  outliers 
# code showing the outliers
```{r}
boxplot.stats(mydata$Area.conventional)$out
```
##code showing the rows with outliers
```{r}
out<-boxplot.stats(mydata$Area.conventional)$out
out_ind<-which(mydata$Area.conventional%in%c(out))
out_ind
```

```{r}
mydata[out_ind,]
```

```{r}
boxplot(mydata$Area.conventional,
        ylab="Production.conventional")
mtext(paste("Outliers:",paste(out,collapse = ",")))
```

```{r}
boxplot.stats(mydata$Area.organic)$out
```

```{r}
boxplot.stats(mydata$Area.total)$out
```
```{r}

```

```{r}
iqr<-IQR(mydata$Production.conventional,na.rm = TRUE)
up<-quantile(mydata$Production.conventional,na.rm = TRUE,0.75)+1.5*iqr#upper range
low<-quantile(mydata$Production.conventional,na.rm = TRUE,0.25)-1.5*iqr #Lower range

```
```{r}
eliminated<-subset(mydata, mydata$Production.conventional>low & mydata$Production.conventional<up)
```

```{r}
Aggregateoutlier<-eliminated%>%
  group_by(Region)%>%
  summarise(across(c(Area.total,Area.conventional,Area.organic,
                     Production.conventional,Production.Organic,Production.total),sum,na.rm=TRUE))
Aggregateoutlier
```


```{r}
newdata<-mydata
newdata$Region<- as.numeric(as.character(newdata$Region))
newdata$Area.conventional<- as.numeric(as.character(newdata$Area.conventional))

```

```{r}
dataout<-newdata[!newdata %in% boxplot.stats(newdata)$out]
```

```